### 2017-02-23 bw-prop.tex
adds BW prop and pagecounting

\documentclass[11pt]{article}

\usepackage[acronym,toc]{glossaries}
\include{acros}
\makeglossaries
\usepackage{fancyhdr}
\usepackage{pagecounting}
\usepackage[dvips]{color}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
% Color Information from - http://www-h.eng.cam.ac.uk/help/tpl/textprocessing/latex_advanced/node13.html

% Trying to bold my name in the bib
\usepackage{xstring}
\def\FormatName#1{%
  \IfSubStr{#1}{Huff}{\textbf{#1}}{#1}%
}

\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\newcommand\bb[1]{\mbox{\em #1}}
\def\baselinestretch{1.1}
%\pagestyle{empty}
\newcommand{\hsp}{\hspace*{\parindent}}
\definecolor{gray}{rgb}{0.4,0.4,0.4}

\newcommand{\authorname}{Kathryn~D.~Huff }
\newcommand{\authoremail}{katyhuff@gmail.com}
\newcommand{\authorsite}{katyhuff.github.com}

\begin{document}

\pagestyle{fancy}
%\pagenumbering{gobble}
%\fancyhead[location]{text}
% Leave Left and Right Header empty.
%\lhead{}
%\rhead{}
\lhead{\textcolor{gray}{\it \authorname}}
\rhead{\textcolor{gray}{\thepage}}
%\rhead{\textcolor{gray}{\thepage/\totalpages{}}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyfoot[C]{\footnotesize \textcolor{gray}{\authorsite}}

\paragraph{Research Goal}

The Advanced Reactors and Fuel Cycles (ARFC) group pursues improved safety and
sustainability of nuclear power. To do so, we model and
simulate the design, safety, and performance of advanced nuclear reactors.
Such reactors  involve tight coupling between physics such as thermal-hydraulic
phenomena, neutron transport, and fuel performance.

The current state of the art in advanced nuclear reactor simulation (e.g. the
CASL DOE innovation hub) is focused on traditional light water reactors.
The ARFC group seeks to extend that state of the art by enabling similarly high fidelity simulation of more advanced reactor designs through development of models and tools for representing unique materials, geometries, and physical phenomena. Current work includes extension
of the MOOSE framework to appropriately model coupled thermal-hydraulics and
neutronics of promising molten salt fuelled reactor designs.
ARFC develops physics kernel extensions and runs simulations using the highly
scalable, fully implicit, Multiphysics Object-Oriented Simulation Environment
(MOOSE) framework from Idaho National Laboratory.

\paragraph{Role of Blue Waters}
Simulations which faithfully capture this coupling at realistic spatial and
temporal resolution are only possible with the aid of high performance
computing resources.  To solve these large systems of partial differential
equations on a finite element mesh in a fully-coupled, implicit way, the MOOSE
framework was designed to take advantage of high performance computing
capability.  The framework conducts fully implicit physics coupling with a
Jacobian Free Newton Krylov (JFNK) solver method and sophisticated automatic
adaptive meshing.  These simulations are memory intensive, so the exceptional
memory capability of the Blue Waters resource will be essential to performant
simulation times. It is also important to note that rendering visualizations of
the results can be computationally intensive was well.

\paragraph{Progress}

While the allocation needs will vary according to the particular simulations, Dr. Huff’s research group will conduct many 2-dimensional and 3-dimensional finite element simulations per year using the MOOSE framework and modules. Such simulations commonly occupy tens of thousands of CPU cores at a time and vary in completion time. The MOOSE framework is shown to scale very well up to 10,000 cores. Transient and multi-scale simulations, which require greater capability per simulation, may occupy up to 100,000 CPU cores at a time. Only a few of those larger simulations are likely to be run per year.

\paragraph{Potential Impact}
Detailed spatially and temporally resolved neutron flux and
temperature distributions can improve designs, help characterize
performance, inform reactor safety margins, and enable validation of numerical
modeling techniques for unique physics.
Future work may include similarly challenging materials and geometries such as
those in sodium cooled, gas cooled, and very high temperature reactor designs
which also boast improved safety and sustainability.

Simulations on the reactor core scale require substantial
computing resources to enable interrogation of
physical phenomena in the complex environments inside a nuclear reactor.

\end{document}