### 2017-03-01 siip.tex
siip draft, so messy

%        File: siip.tex
%     Created: Mon Feb 27 08:00 AM 2017 C
% Last Change: Mon Feb 27 08:00 AM 2017 C
%
\documentclass[11pt]{article}

\usepackage[acronym,toc]{glossaries}
\include{acros}
\makeglossaries
\usepackage{fancyhdr}
\usepackage{pagecounting}
\usepackage[dvips]{color}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

% Trying to bold my name in the bib
\usepackage{xstring}
\def\FormatName#1{%
          \IfSubStr{#1}{Huff}{\textbf{#1}}{#1}%
          }

          \usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
          \newcommand\bb[1]{\mbox{\em #1}}
          \def\baselinestretch{1.1}
          %\pagestyle{empty}
          \newcommand{\hsp}{\hspace*{\parindent}}
          \definecolor{gray}{rgb}{0.4,0.4,0.4}

          \newcommand{\authorname}{Kathryn~D.~Huff }
          \newcommand{\authoremail}{katyhuff@illinois.edu}
          \newcommand{\authorsite}{arfc.npre.illinois.edu}

          \begin{document}

          \pagestyle{fancy}
          %\pagenumbering{gobble}
          %\fancyhead[location]{text}
          % Leave Left and Right Header empty.
          %\lhead{}
          %\rhead{}
          \lhead{\textcolor{gray}{Investigator: Prof. \authorname\\\authoremail}}
          \rhead{\textcolor{gray}{Advanced Reactors and Fuel Cycles\\Dept. of Nuclear, 
          Plasma, and Radiological Engineering}}
          %\rhead{\textcolor{gray}{\thepage/\totalpages{}}}
          \renewcommand{\headrulewidth}{0pt}
          \renewcommand{\footrulewidth}{0pt}
          \fancyfoot[C]{\footnotesize \textcolor{gray}{\authorsite}}

          \section{The Problem}
          As you read this, thousands of professors in the United States are 
          simultaneously preparing lessons on solving a conduction equation. 
          They are doing so largely without recieving feedback from one another 
          or directly building on one another's experience 
          \cite{wilson_software_2014}.  It's distressing but true that 
          professors spend 10-40\% of their time each year repeating curriculum 
          development efforts already tackled by numerous colleagues within 
          their disciplines. What's worse is that these efforts are rarely, if 
          ever, reviewed by, shared with, or extended upon by others.

          \section{The Proposed Solution}
          We propose a proof-of-concept for collaborative, open source, 
          curriculum development that will break down barriers between 
          classrooms and improve the transfer of lessons learned across universities.

          The goal will be to explore whether curriculum development can 
          operate like open source software development does, drawing on a 
          community of experts toward creating a shared core resource. In open 
          source software, developers share code revisions in online 
          repositories. In a typical project, the main ``fork'' holds the 
          official copy of the software package. Developers each have their own 
          forks, copies of the repository, where they can work on features and 
          bug fixes. When those features and bug fixes are ready for prime 
          time, the developer makes a ``pull request'' to the main fork, it is 
          reviewed by their collaborators, and it is eventually merged into the 
          main fork where it can be used by all.

          \begin{figure}
                  \centering
                  \begin{subfigure}{.4\textwidth}
                            \centering
                              \includegraphics[width=.8\linewidth]{git-flow}
        \caption{This figure, from \cite{scopatz_effective_2016}, captures a process, often called Git Flow, through which a new feature or bug fix enters a piece of open source software \cite{scopatz_effective_2016}.}
                                  \label{fig:sub1}
                  \end{subfigure}\hfill%
                  \begin{subfigure}{.4\textwidth}
                            \centering
                              \includegraphics[width=.8\linewidth]{siip-flow}
  \caption{This adaptation imagines the process in the context of learning module development, using the same git version control system and GitHub repository hosting framework.}
                                  \label{fig:sub2}
                  \end{subfigure}
                  \caption{A figure with two subfigures}
                  \label{fig:test}
          \end{figure}

          This initiative already involves a few peer professors. Six faculty 
          in Nuclear Engineering from UC Berkeley, the University of Wisconsin 
          at Madison, Kansas State, U. South Carolina, and U. 
          Tennessee-Knoxville are involved. We all use GitHub \cite{github} to 
          store, revise, and collaborate on research, especially source code. A 
          few of us have already started to host our course curricula online as 
          well, but these are typically single author repositories 
          \cite{skutnik,wilson,slaybaugh,scopatz,borelli,roberts}.

          The participants will collaborate on a master set of learning 
          modules for an upper-division course in nuclear engineering : 
          The Nuclear Fuel Cycle.  The curriculum will be hosted on GitHub, 
          used by all of us, and improved continually as we learn and grow as 
          instructors.

          \paragraph{Each learning module} may include active in-class exercises, 
          presentation notes, homework problems, project descriptions, and 
          evaluation tools.

          However, curriculum is not the same as source code, because each 
          instructor wants and deserves to impart their own style of teaching 
          (and corny jokes) on their instruction. This challenge has been 
          encountered by collaborative instruction organizations such as 
          Software Carpentry, in which forks of the original material diverge 
          without any strong incentive for contribution back to the master 
          material \cite{wilson_software_2014}.

          This proposal suggests that fine-grained modularity and a clear 
          dependency graph may assist in overcoming this goal. Further, a small 
          and devoted group of like-minded nuclear engineering professors 
          should be able to compromise on a cohesive set of modular materials 
          that can be mixed and matched according to individual preference. 
          This work could provide an important proof of concept for groups of 
          instructors willing to collaborate on open source curriculum for:

          - core courses with many sections in a single university
          - niche courses taught by a select group of professors across 
          universities
          - fundamental courses in small fields (like nuclear engineering)

          Why do educators not diff, patch, and merge one another's . 
          Collaboratively construct and improve lesson materials.

          Scientists are more than happy to build upon one another's work, form 
          collaborations with others in their field. But, 
          when it comes to educating, where is the sharing of lessons learned 
          and collaboration?

          \section{The Potential Impact}
          \cite{wilson_software_2014}.

          Doing code review at the end of the work isn't useful. What works is 
          incremental code review. This proposal suspects that the same is true 
          for curriculum review. 
          \cite{wilson_software_2014}.

          Education is an inherently distributed system, but this need not be a 
          hinderance to collaboration.

          \bibliographystyle{katyunsrt}
          \bibliography{bibliography}

          \section{Departmental Support}
          The Department of Nuclear, Plasma, and Radiological Engineering will 
          support this work with relese time through startup funds. 
          Additionally, NPRE will support workshop activities through 
          administrative effort and by providing space.
          \end{document}