### 2017-03-02 2017-03-02-ae3.tex
final draft

%        File: 2017-03-02-ae3.tex
%     Created: Thu Mar 02 09:00 AM 2017 C
% Last Change: Thu Mar 02 09:00 AM 2017 C
%
\documentclass[11pt]{article}

\usepackage[acronym,toc]{glossaries}
\include{acros}
\makeglossaries
\usepackage{fancyhdr}
\usepackage{pagecounting}
\usepackage[dvips]{color}
\usepackage{graphicx}

% Trying to bold my name in the bib
\usepackage{xstring}
\def\FormatName#1{%
          \IfSubStr{#1}{Huff}{\textbf{#1}}{#1}%
          }

          \usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
          \newcommand\bb[1]{\mbox{\em #1}}
          \def\baselinestretch{1.1}
          %\pagestyle{empty}
          \newcommand{\hsp}{\hspace*{\parindent}}
          \definecolor{gray}{rgb}{0.4,0.4,0.4}

          \newcommand{\authorname}{Kathryn~D.~Huff }
          \newcommand{\authoremail}{katyhuff@illinois.edu}
          \newcommand{\authorsite}{arfc.npre.illinois.edu}

          \begin{document}
\pagestyle{fancy}
          %\pagenumbering{gobble}
          %\fancyhead[location]{text}
          % Leave Left and Right Header empty.
          %\lhead{}
          %\rhead{}
          \lhead{\textcolor{gray}{Prof. \authorname\\\authoremail}}
          \rhead{\textcolor{gray}{Academy for Excellence in Engineering 
          Education}}
          %\rhead{\textcolor{gray}{\thepage/\totalpages{}}}
          \renewcommand{\headrulewidth}{0pt}
          \renewcommand{\footrulewidth}{0pt}
          \fancyfoot[C]{\footnotesize \textcolor{gray}{\authorsite}}

          \begin{itemize}
                  \item Hello everyone. ``What's the cube root of eight?''
                  \item ``Anyone? Yes, you, please what's the cube root of eight?'' 
                  \item Yes, it's 2. Thank you. 
          \end{itemize}

          \paragraph{Now what just happened?}

          All of you knew the answer, but not all of you raised your hands.
          \paragraph{Some of you did,} because you felt comfortable, confident, 
          and engaged. 
          \paragraph{Some of you didn't.} Perhaps you were checking your email 
          and I caught you offguard.
          \paragraph{According to research on stereotype threat} some of you 
          may have resisted raising your hands because you were under threat of 
          confirming a negative stereotype about your social or demographic group. 
          In the case of academics during a talk, perhaps the fear is that you 
          would appear to be an attention-seeking know-it-all. Of course, in 
          the case of an engineering classroom, there are much more sinister 
          stereotypes which threaten the comfort of underrepresented minority 
          and female students in particular.

          \paragraph{Ok, fine, not everyone raised their hands.} But then what 
          happened?

          \paragraph{I picked that guy. But why?} Maybe because I'm right 
          handed. In the case of a classroom after many sessions with the same 
          students, perhaps I'll choose him because he reliably gives 
          the correct answer.
          I \emph{could} conciously try to avoid that implicit bias by  
          explicitly choosing someone who didn't their hand. However, 
          that solution isn't fair either and can feel like punishment.

          \paragraph{Ok that was a fun experiment which I think 
          demonstrates a problem that is already very well understood by anyone 
          who has ever been in a classroom.}

          I've tried to level the playing field in my class when I ask a 
          question, by randomizing the name that I call on.

          While it's initially uncomfortable, it does help me learn everyone's 
          names, it keeps folks on their toes, it acts as a spot check on 
          attendance.  I'll sample from the presenter's names as an example.

          \begin{itemize}
                  \item What's the square root of nine?
          \end{itemize} 
          It also has the added benefit of showing that even the really smart kid doesn't always know the answer.
          \begin{itemize}
                  \item What's the prompt energy of a fission neutron?
          \end{itemize}

          \bibliographystyle{unsrt}
          \bibliography{bibliography}

          \end{document}