### 2017-07-09 msbr_reproc.tex
rephrasing in online reprocessing method section

\documentclass{anstrans}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Online reprocessing simulation for thorium-fueled molten salt breeder 
reactor}

\author{Andrei Rykhlevskii, Alexander Lindsay, Kathryn Huff}

\institute{
        Department of Nuclear, Plasma, and Radiological Engineering, University 
        of Illinois at Urbana-Champaign \break
        Urbana, IL
}

\email{andreir2@illinois.edu}

%%%% packages and definitions (optional)
\usepackage{graphicx} % allows inclusion of graphics
\usepackage{caption}  % allows center figures caption
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF
\usepackage[section]{placeins}

\usepackage[acronym,toc]{glossaries}  % acronyms inclusion
\include{acros}

\usepackage{tabularx} % for nice wrapping of table text

\makeglossaries

\graphicspath{{figures/}}

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\grad}{\vec{\nabla}} % gradient
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
The thermal spectrum \gls{MSR} is an advanced type of reactor that consists of 
constantly circulating liquid fuel (i.e., mixture of $LiF-BeF_2-ThF_4-UF_4$ or 
$LiF-BeF_2-ZrF_4-UF_4$) and solid graphite moderator structures. This liquid 
fuel form leads to immediate advantages over traditional, 
solid-fueled, reactors. The molten-salt carrier with dissolved 
fissile and/or fertile material allows for online refuelling and 
reprocessing. Thus, \glspl{MSR} can potentially operate years without shutdown, 
achieving maximum fuel utilization and outstanding neutron economy 
\cite{leblanc_molten_2010}. Moreover, this type of fuel does not need 
complex fabrication since the fissile material could be transported from any 
enrichment plant to the \gls{NPP} in the form 
of uranium hexafluoride ($UF_6$).  Additionally, \glspl{MSR} have a high level 
of inherent safety due to their characteristically 
strong negative temperature coefficient of reactivity, near-atmospheric 
pressure in the primary loop, stable coolant, passive decay heat cooling, and 
small excess reactivity \cite{elsheikh_safety_2013}.

The thorium-fueled \gls{MSBR} was developed in the early 1970s by \gls{ORNL} 
specifically to realize the promise of the thorium fuel cycle which allows the 
use of natural thorium instead of enriched uranium as the fertile element. 
Thorium breeds the fissile $^{233}U$ and avoids uranium enrichment 
\cite{robertson_conceptual_1971}. In the matter of nuclear fuel cycle, the 
thorium cycle produces a reduced quantity of plutonium and \glspl{MA} 
compared to the traditional uranium fuel cycle. Consequently, it may 
significantly increase proliferation resistance when the \gls{MSR} operates in the 
breeder regime. The \glspl{MSR} also could be employed as a converter reactor for 
transmutation or spent fuel from current \glspl{LWR}.

Recently, interest in \glspl{MSR} has resurged, with multiple new companies 
pursuing commercialization of \gls{MSR} designs\footnote{Examples include both 
liquid-fueled molten salt designs from Transatomic, Terrapower, Terrestrial, 
and Thorcon.}.
To further develop these \gls{MSR} concepts, particularly with respect to their  
strategies for online reprocessing and refueling, computational analysis methods capturing
their unique reactor physics and process chemistry are needed.
However, most contemporary nuclear reactor physics software is unable to 
perform depletion calculations in an online reprocessing 
regime. Powers \emph{et al.} suggested a novel method for conducting 
depletion simulations for \gls{MSR}. This suggested method takes into account 
fuel salt composition changes due to online 
reprocessing and refueling based on the deterministic computer code NEWT in 
SCALE \cite{powers_new_2013}. This approach was later used by Jeong \emph{et 
al.} to find an equilibrium fuel composition for the \gls{MSBR} and was 
validated with an \gls{MCNP}/CINDER90 model \cite{jeong_equilibrium_2016}.

The current paper presents a single-cell model developed using the
continuous-energy Serpent 2 Monte Carlo reactor physics software. It was 
employed to establish a Serpent-based method for finding the equilibrium core composition and 
core depletion of the \gls{MSBR}.
All calculations presented in this paper were performed using the Serpent 2 
code version 2.1.29 with ENDF/B-VII.0
nuclear data \cite{leppanen_serpent_2012,chadwick_endf/b-vii.0:_2006}. Serpent 
2 is an improvement upon Serpent 1, and contains a complete 
redesign of memory management using hybrid OpenMP + MPI parallelization.  This 
hybrid parallelization is important in depletion calculations using computer 
clusters with multiple cores \cite{leppanen_serpent_2015}. This work used the 
built-in Serpent 2 depletion capabilities and its built-in online reprocessing 
subroutine. Another feature of the \gls{MSBR}, its circulating liquid fuel and 
corresponding delayed neutron precursor drift, is not treated here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Description of the actual work}
The \gls{MSBR} is a thermal spectrum reactor. The reactor vessel has a diameter of 
$680 cm$ and a height of $610 cm$. It contains a molten fluoride fuel-salt mixture 
which performs two functions: to generate heat in the moderated region and to 
transport heat energy from the core to primary heat exchanger using the primary 
salt pump. The vessel also contains graphite blocks for neutron moderation and 
reflection. The lithium in the fuel-salt solution is enriched to 99.995\% 
$^7$Li because $^6$Li is a very strong neutron poison and becomes tritium upon 
neutron bombardment. In this study, the 0.005\% atomic fraction of $^6$Li has 
been taken into account because even such a small amount of isotope with very 
high absorption cross section can significantly impact the neutron flux energy 
distribution and, consequently, depletion calculation results. 
Table~\ref{tab:data} is the summary of the major \gls{MSBR} parameters used by 
this model \cite{robertson_conceptual_1971}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup[table]{
        labelsep = newline,
        name = TABLE, justification=justified,
        singlelinecheck=false,%%%%%%% a single line is centered by default
        labelsep=colon,%%%%%%
skip = \medskipamount}
\begin{table}[h!]
        %\centering
        %\begin{tabular}{p{0.46\linewidth} p{0.50\linewidth}} \toprule
        \begin{tabularx}{\linewidth}{l X} \toprule 
                Thermal capacity of reactor           & 2250 MW(t)
                \\ \midrule
                %Net electrical output                 & 1000 MW(e) \\ \midrule
                %Net thermal efficiency        & 44.4\%
                %\\ \midrule
                Salt volume fraction in central core zone     & 0.132
                \\ \midrule
                Salt volume fraction in outer core zone       & 0.37
                \\ \midrule
                Fuel-salt inventory (Zone I)                  & 8.2 m$^3$	
                \\ \midrule
                Fuel-salt inventory (Zone II)                 & 10.8 m$^3$	
                \\ \midrule
                Fuel salt components                  & 
                LiF-BeF$_2$-ThF$_4$-$^{233}$UF$_4$-$^{239}$PuF$_3$	\\ 
                \midrule
                Fuel salt composition                 & 
                71.767-16-12-0.232-0.0006 mole\%
                \\
                \bottomrule
        \end{tabularx}
        \caption{Summary of principal data for MSBR.}
        \label{tab:data}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The \gls{MSBR} core consists of two different zones. The central 
zone, Zone I, in which 13.2\% of the volume is fuel salt and 86.8\% graphite. 
The first zone is composed of 1320 graphite cells, 2 graphite control rods, and 
2 safety rods consisting of boron carbide clad. The under-moderated zone, Zone 
II, with 37\% fuel salt, and radial reflector, surrounding the more active 
portion serves to diminish neutron leakage from the reactor core. In this work, 
only a Zone I element with a $3.42138 cm$ diameter central fuel channel and molten-salt 
fuel volume fraction of 0.132 was considered. Figure~\ref{fig:zoneI} shows the 
geometry of the unit cell model. The boundary condition of the unit cell is 
periodic. The density of the fuel salt is $3.3304 g/cm^3$ and the graphite density is 
$1.843 g/cm^3$. The temperature of fuel and graphite is fixed at 908K for the whole 
reprocessing cycle\cite{robertson_conceptual_1971}.

\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=0.90\linewidth]{zone_I_mesh.png}
        \caption{MSBR unit cell of Zone I geometry.}
        \label{fig:zoneI}
\end{figure}

\FloatBarrier
\subsection{Online reprocessing method}

Currently, researchers 
investigating the \glspl{MSR} 
typically develop custom supporting tools and scripts to simulate 
online reprocessing and refueling. Most utilize stochastic (i.e. 
\gls{MCNP}) or deterministic (i.e. SCALE) methods for depletion and supplement 
these with custom Python scripts to capture the impacts of reprocessing 
\cite{jeong_equilibrium_2016,park_whole_2015}. Serpent 2 is the 
first commonly used monte carlo software to support continuous material 
reprocessing. It does so by allowing the user to 
create multiple material flows into and out of the fuel.

The \gls{MSBR} has the capability to remove every 20 seconds all poisons (e.g. 
$^{135}$Xe), noble metals and gases (e.g. $^{75}$Se, $^{85}$Kr) which also have 
relatively high absorption cross section. The thorium-232 absorbs thermal 
neutrons and produce $^{233}$Pa which then decay into the fissile $^{233}$U. 
The problem of protactinium is a large absorption cross section in the thermal 
energy spectrum, therefore $^{233}$Pa continuously removing from fuel salt into 
protactinium decay tank and allowing $^{233}$Pa to decay to $^{233}$U without 
reactor poisoning. The reactor reprocessing system designed to separate 
$^{233}$Pa from the molten-salt fuel over 3 days, held it to produce $^{233}$U 
and return it back to the primary loop. This feature allows to keep the 
neutrons losses to protactinium and fission products to a very low level, and 
increases the efficiency of $^{233}$U breeding \cite{robertson_conceptual_1971}.

A major problem with the reprocessing method is that different nuclides have 
specific removal rates. On the one hand, if the depletion time intervals are 
very short it means specific isotope being instantly removed, and it is hard to 
obtain equilibrium composition because enormous large number of depletion 
steps. On the other hand, if the depletion calculation time interval is too 
long the results would not truly represents \gls{MSBR} conceptual design. 
Following this idea, the time interval for depletion calculations was selected 
3 days that correlate with the removal interval of $^{233}$Pa which is the 
major isotope for producing $^{233}$U. Thorium is continuously adding to 
maintain the initial mass fraction of $^{232}$ThF$_4$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}

Using the methodology described previously, the \gls{MSBR} unit cell depletion 
analysis was performed to find equilibrium core conditions. In this section, 
the calculation results including multiplication factor, neutron flux energy 
spectrum, atomic density of major isotopes are summarised.

\subsection{Equilibrium state analysis}

The method approximate a single representative unit cell for the whole 
\gls{MSBR} core, consequently, it does not take into consideration different 
fuel-moderator volume ratio for Zone I, Zone II, annulus, and reflector. The 
initial multiplication factor during depletion calculation is selected for 
state with fully withdrawn control and poison rods position which gives high 
excess reactivity in the beginning of the cycle (approximately 5000 pcm). 
Standard deviation is about 100 pcm.

\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.03\linewidth]{keff.png}
        \caption{Infinite multiplication factor during 4 years depletion 
        simulation.}
        \label{fig:keff}
\end{figure}

\FloatBarrier

Figure~\ref{fig:keff} shows the infinite multiplication factor for 4-year 
reprocessing cycle calculated by Serpent 2 with ENDF/B-VII.0 nuclear data. The 
high standard deviation causes the multiplication factor fluctuations. 
Protactinium-233 continuously removing into the tank for protactinium decay. 
$^{233}$Pa has a half-life of 27 days and beta decays into $^{233}$U which as 
fresh fuel goes back to the reactor core. The infinite multiplication factor 
decreases first 400 days of depletion due to strong absorbers (e.g. $^{233}$Th, 
$^{234}$U) accumulation which causes relatively high fuel ($^{233}$U) refill 
inflow to keep reactor critical. During reactor operation producing fissile 
materials other than $^{233}$U in the core (e.g. $^{235}$U, $^{239}$Pu) which 
makes it possible to decrease fresh fuel refill rate after ~1 year of 
operation.

\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.03\linewidth]{fuel_composition.png}
        \caption{Normalized number density of major isotopes during 4 years 
        depletion.}
        \label{fig:compos}
\end{figure}

\FloatBarrier

The analysis of the fuel salt composition variation gives clearer information 
about equilibrium core state. Figure~\ref{fig:compos} shows the normalized 
number density of important isotopes that have strong influence on the core 
neutronics at the beginning of each depletion time interval. The number density 
of protactinium is very low (less than $10^{16}$1/cm$^3$) but some small amount 
of it produced during 3-days reprocessing step. It can be seen that the 
multiplication factor reaches the equilibrium state at about 950  days. 
Figure~\ref{fig:rates} represents the rates of materials in online reprocessing 
flows during 4 years of depletion calculations. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\captionsetup[table]{
%  labelsep = newline,
%  name = TABLE, justification=justified,
%  singlelinecheck=false,%%%%%%% a single line is centered by default
%  labelsep=colon,%%%%%%
%  skip = \medskipamount}
%\begin{table}[h!]
%\centering
%\begin{tabular}{p{0.25\linewidth} p{0.3\linewidth} p{0.3\linewidth}} \toprule
%      Isotope & First 400 days     & After 400 days          
%\\ \midrule
%$^{232}$Th  & 3.704$\times10^{12}$ & 3.086$\times10^{12}$
%\\ \midrule
%$^{233}$Pa  & 3.704$\times10^{12}$ & 3.086$\times10^{12}$
%\\ \midrule
%$^{233}$U  & $1.05134\pm0.00116$ & $1.00736$
%\\
%\bottomrule
%\end{tabular}
%  \caption{Materials flows rates during \gls{MSBR} online reprocessing 
%  [1/s*cm$^3$].}
%  \label{tab:rates}
%\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.03\linewidth]{rates_fuel.png}
        \caption{Materials flows rates during online reprocessing.}
        \label{fig:rates}
\end{figure}
From the graph below we can see that to keep reactor critical higher $^{233}$U 
flow rate required for first 400 days of operation and after that much smaller 
uranium-233 flow rate from the protactinium decay tank needed. The $^{232}$Th 
rate slightly decreases over 4 years of operation due to other than $^{233}$U 
fissile materials accumulation.

\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.03\linewidth]{rates_outflow.png}
        \caption{Materials flows rates for the protactinium decay tank during 
        \gls{MSBR} online reprocessing.}
        \label{fig:outflow}
\end{figure}
As shown in Figure~\ref{fig:outflow}, the tank for protactinium decay 
accumulates $^{233}$Pa approximately 200 days and fresh fuel ($^{233}$U) flow 
established also after 200 days. Uranium produced in the tank by protactinium 
decay is separated by circulation of the salt through a flourinator. The fully 
processed molten salt, on its way back to the primary loop, has uranium added 
at the rate required to maintain or adjust the fissile material concentration 
and, hence, the reactivity, in the reactor core as desired.

\subsection{Neutron spectrum}

Figure~\ref{fig:spectrum} represents the normalized neutron flux distribution 
of the initial and the equilibrium core compositions. The spectrum for the 
equilibrium state is harder than the initial state due to heavy fission 
products.  
\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.04\linewidth]{spectrum.png}
        \caption{Normalized neutron energy spectrum for initial and equilibrium 
        composition.}
        \label{fig:spectrum}
\end{figure}

\FloatBarrier

In this work only Zone I unit cell with 0.132 volume fraction of fuel 
considered, neutron spectrum for Zone II cell where salt fraction is 37\% 
should be harder and the peak in thermal energy region predicted to be much 
lower. To obtain high fidelity neutron energy spectrum whole \gls{MSBR} core 
analysis is required.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
The depletion calculation of the \gls{MSBR} unit cell model for finding the 
equilibrium states was performed using the Serpent 2 Monte Carlo code to 
simulate simplified case of the online reprocessing and refueling to find 
equilibrium material composition. When running depletion calculation, the 
fission products are removed and fertile/fissile materials are added to fuel 
salt every 3 days. The important MSR feature, online reprocessing I\& refueling 
is implemented in the Serpent 2 material burnup routine. The results of this 
study indicate that from the depletion calculation the multiplication factor 
slowly decreases and reaches to the equilibrium state. The most obvious finding 
to emerge from the analysis of initial and equilibrium materials composition is 
that neutron energy spectrum is harder for equilibrium state because 
significant amount of heavy fission products were accumulated in the \gls{MSBR} 
core.

These results are contrary with those of Jeong and Park (2016) who suggested 
two different unit cell models and uses \gls{MCNP} with Python-script to 
simulate \gls{MSBR} online reprocessing to find equilibrium composition. This 
inconsistency may be due to different fuel fraction in the unit cell (Jeong and 
Park selected 20.6\% salt fraction). To obtain better results in online 
reprocessing simulation number of future efforts expected. First, depletion 
simulation will be performed using full-core tree-dimensional high-fidelity 
model of \gls{MSBR} developed in Serpent 2. In this case, different 
fuel-moderator volume ratios for the reactor Zone I, Zone II, annulus, 
reflectors would be taken into account to find accurate multiplication factor, 
neutron spectrum and, hence, depleted composition. Secondly, additional Serpent 
2 flow control system subroutine should be developed to simulate adjusting 
material flows (e.g. rate of removing $^{233}$Pa from the salt and adding 
fissile $^{233}$U from the tank for protactinium decay) depending upon current 
reactivity value which is more promising reactivity control method than moving 
control and safety rods. Finally, temperature effect of reactivity for both 
fuel salt and graphite should be calculated to find optimal effective 
multiplication factor range.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ans}
\bibliography{bibliography}
\end{document}