### 2017-08-31 results.tex
shortened page to 4

\section{Results}

\subsection{Historical Operation of \gls{EU} Reactors}

\begin{table}[h]
	\centering
	\scalebox{0.77}{
		\begin{tabular}{|c|c|c|c|}
			\hline
			Category & Unit & Value & Specifics\\ \hline
			Total UOX Usage & MTHM & 181,471 &  \\ \hline
			Total MOX Usage & MTHM & 6,302 & \\ \hline
			Total Used UOX Stored & MTHM & 141,659 & \gls{UNF} that is not reprocessed\\ \hline
			Total Used  MOX Stored & MTHM & 3,611 & \gls{UNF} that is not reprocessed \\ \hline
			Total Tailings & MTHM & 1,081,826 & \\ \hline
			Total Natural U Used & MTHM & 1,269,897 & \\ \hline
		\end{tabular}

\Cref{tab:sim_result} lists the important metrics
obtained from the first simulation. The following
values are the \gls{EU} inventory and history at year 2050,
and will be reprocessed in the second simulation.

}
	\caption{Simulation Results for Historical Nuclear Operation of \gls{EU} Nations}
	\label{tab:sim_result}
	\end {table}
	Figures \ref{fig:eu_tail} and \ref{fig:eu_snf} show the 
	timeseries of mass of tailings and used fuel accumulation in \gls{EU}.
	Figure \ref{fig:eu_fuel} shows the amount of fuel used in \gls{EU}.

	\begin{figure}[htbp!]
		\begin{center}
			\includegraphics[width=\columnwidth]{./images/eu_future/tails.png}
		\end{center}
		\caption{Timeseries of Tails Mass in the \gls{EU}.}
		\label{fig:eu_tail}
	\end{figure}

	\begin{figure}[htbp!]
		\begin{center}
			\includegraphics[width=\columnwidth]{./images/eu_future/total_fuel.png}
		\end{center}
		\caption{Timeseries of Total Fuel Usage in \gls{EU}.}
		\label{fig:eu_fuel}
	\end{figure}

	\begin{figure}[htbp!]
		\begin{center}
			\includegraphics[width=\columnwidth]{./images/eu_future/snf_discharge.png}
		\end{center}
		\caption{Timeseries of Used Nuclear Fuel in \gls{EU}.}
		\label{fig:eu_snf}
	\end{figure}
	\FloatBarrier

	\subsection{French \gls{SFR} Transition Scenario}

	Reprocessing \gls{UNF} collected from all EU nations can start approximately
	270 \glspl{SFR}, which is more than enough for two generations of 66GWe \gls{SFR}
	fleet. With the \gls{SFR} breeding ratio of over one, France can transition into
	a fully \gls{SFR} fleet without extra construction of \glspl{LWR}.

	From Varaine et al. \cite{varaine_pre-conceptual_2012}, a French
	ASTRID-type \gls{SFR} of capacity 600 MWe needs $1.225$ tons of
	plutonium a year, with an initial plutonium loading of $4.9$ tons. 
	Thus, the number of \glspl{SFR} that can be loaded with the reprocessed
	plutonium from \gls{UNF} can be estimated to 270 \glspl{SFR},
	assuming infinite reprocessing and fabrication capacity as well as
	abundant depleted uranium supply.

	Also, assuming that \gls{MOX} can be recycled indefinitely,
	used \gls{MOX} from an ASTRID reactor contains enough plutonium to produce a \gls{MOX} fuel with
	the same mass, if mixed with depleted uranium. For example,
	used \gls{MOX} from an ASTRID reactor is assumed to be 12.6\% plutonium
	in this simulation, whereas a fresh \gls{MOX} is 11\% plutonium.
	Separating plutonium from used \gls{MOX} from
	an ASTRID reactor can create \gls{MOX} of the mass of used \gls{MOX}.
	The plutonium breeding ratio in this simulation is thus assumed to be
	$\approx 1.145$.

	\begin{figure}[htbp!]
		\begin{center}
			\includegraphics[width=\columnwidth]{./images/french-transition/where_fuel.png}
		\end{center}
		\caption{Timeseries of fuel loaded into \glspl{SFR}}
		\label{fig:fuel}
	\end{figure}

	\begin{figure}[htbp!]
		\begin{center}
			\includegraphics[width=\columnwidth]{./images/french-transition/pu.png}
		\end{center}
		\caption{Separated plutonium discharge from Reprocessing Plant}
		\label{fig:pu_no_cum}
	\end{figure}

	\Cref{fig:fuel} shows \gls{MOX} loaded in the \glspl{SFR} per month.
	The spikes are due to initial fuel demand for new deployment of \glspl{SFR}.
	The initial loading of new \glspl{SFR} are done with the \gls{MOX} created
	from legacy \gls{UNF}. Once there are enough amounts of extra plutonium creation
	by deployed \glspl{SFR}, the legacy \gls{UNF} is no longer used. 
	\begin{table}[h]
		\centering
		\scalebox{0.86}{
			\begin{tabular}{|c|c|c|}
				\hline
				Category & Unit & Value  \\ \hline
				Total MOX used & MTHM & 127,640  \\ \hline
				Total \glspl{SFR} Deployed & & 220 \\ \hline
				Total Plutonium Reprocessed & MTHM & 16,352 \\ \hline
				Total MOX from UOX Waste & MTHM & 6,570  \\ \hline
				Total MOX from MOX Waste & MTHM  & 121,070 \\ \hline
				Total Tails used & MTHM & 116,153 \\ \hline
				Total legacy UNF reprocessed & MTHM & 77,082 \\ \hline
				Total Reprocessed Uranium Stockpile & MTHM & 184,172 \\ \hline
				Total Reprocess Waste & MTHM & 16,352 \\ \hline
			\end{tabular}}
			\caption {\gls{SFR} Simulation Results}
			\label{tab:sfr_sim_result}
			\end {table}

	\Cref{fig:pu_no_cum} shows the separated plutonium discharge
	per month from the reprocessing plant. The plutonium outflux
	does not precisely follow the fuel demand because \Cyclus agents have
	material buffers that store commodity fuel for later usage. 
	\Cref{tab:sfr_sim_result} lists metrics obtained from the second simulation.